---
title: "GDP"
author: "Yi Zheng"
date: "October 11, 2015"
output: html_document
---

This is the R markdown document for answering questions on Gross Domestic Product data for the 190 ranked countries. 

Required packages: curl. Install and/or load these packages before trying the code below.
```{r}
library(curl)
```

##Read in and clean GDP data
```{r}
fileurl<-'https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv'
#We want the data start from the fifth line and we only want the first, third, fourth and fifth columns.
gdp<-read.csv(curl(fileurl),skip=4,na.strings = c("..","Not available.","..Not available"),stringsAsFactors = F)
gdp<-dplyr::select(gdp,X,X.1,X.3,X.4)
names(gdp)<-c('CountryCode','Ranking','Economy','GDP')
#Change the format of some of the variables and be careful with the number containing period
gdp$Ranking<-as.numeric(gdp$Ranking)
gdp$GDP<-as.numeric(gsub(",","", gdp$GDP))
sum(is.na(gdp$Ranking))
#Delete the observations which have ranking blank
gdp1<-gdp[1:(nrow(gdp)-sum(is.na(gdp$Ranking))),]
sum(is.na(gdp1$Ranking))
sum(is.na(gdp1$CountryCode))
sum(is.na(gdp1$Economy))
sum(is.na(gdp1$GDP))
```

##Read in and clean country data. Noticing that we only want CountryCode and Income.Group column. 
```{r}
fileurl2<-'https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv'
country<-read.csv(curl(fileurl2),header = T,stringsAsFactors = F)[,c(1,3)]
nrow(subset(country,country$Income.Group==''))
#24 countries with no Income.Group
sum(is.na(country$Income.Group))
sum(is.na(country$CountryCode))
```

##1. Match the data based on the country shortcode. How many of the IDs match? 
```{r}
gdp2<-merge(gdp1,country,by='CountryCode')
length(unique(gdp2$CountryCode))
```
###Therefore, 189 of 190 obervations have matched IDs.

##2. Sort the data frame in descending order by GDP rank (so United States is last). What is the 13th country in the resulting data frame?
```{r}
gdp3<-gdp2[order(gdp2$GDP,decreasing = F),]
gdp3[13,]
```
###The 13th country is St. Kitts and Nevis.

##3. What are the average GDP rankings for the "High income: OECD" and "High income: nonOECD" groups? 
```{r}
gdpOECD<-subset(gdp3,gdp3$Income.Group=='High income: OECD')
mean(gdpOECD$Ranking)
gdpnonOECD<-subset(gdp3,gdp3$Income.Group=='High income: nonOECD')
mean(gdpnonOECD$Ranking)
```
###The average GDP rankings for the "High income: OCED" group is 32.96667 and that for the "High income: nonOECD" group is 91.91304.


##4. Cut the GDP ranking into 5 separate quantile groups. Make a table versus Income.Group. How many countries are Lower middle income but among the 38 nations with highest GDP?
```{r}
cutpoint <- quantile(gdp3$Ranking, probs=seq(0, 1, 0.2),na.rm=T)
gdp3$Rank.Group<-cut(gdp3$Ranking,cutpoint)
levels(gdp3$Rank.Group)<-c(1,2,3,4,5)
#5 stands for the lowest level and 1 stands for the highest level
table(gdp3$Income.Group,gdp3$Rank.Group)
sum(is.na(gdp3$Rank.Group))
```
###5 countries are Lower middle income but among the 38 nations with highest GDP.